Thesis Title

### 2022-01-28 chapters/introduction.tex
Progress on introduction

\section{Genomics and metagenomics in clinical microbiology}

@@

>>>>

\section{The global impact of microbial pathogens}

The Global Burden of Disease (GBD) 2019 study reported that microbial pathogens are responsible for more than 400 million years of life lost annually across the globe, a higher burden than either cancer or cardiovascular disease \citep{vos_global_2020}. In particular, lower respiratory infections, diarrhoeal diseases, HIV/AIDS and tuberculosis were amongst the five leading causes of of global total years of life lost. More recently, the COVID-19 pandemic, declared as such by the World Health Organization (WHO) on 11 March 2020 after the emergence and global spread of the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2), and as of January 2022 has cause more than 5.63 million deaths worldwide \citep{ritchie_coronavirus_2020}, making it one of the deadliest pandemics in history. Coronavirus have been responsible for three of the eighteen major pandemics registered throughout modern history \citep{piret_pandemics_2021}, all occurring after the year 2000. \textit{Yersinia pestis}, responsible for three pandemics of plague, \textit{Vibrio cholerae}, with seven cholera pandemics, and Influenza A virus, the causative agent of five flu pandemics, are responsible for the remaining ones, with Influenza being the only other pathogen with a pandemic registered after 2000. Recent decades have also witnessed the emergence of additional virulent pathogens, including the Ebola virus, West Nile virus, Dengue virus and Zika virus, particularly in lower-income countries.

In addition to the emergence of virulent pathogens, the rise of antimicrobial resistance (AMR) poses a major threat to human health around the world. In 2019 there were an estimated 4·95 million deaths associated with bacterial AMR \citep{murray_global_2022}. In 2017, the WHO released The Global Priority Pathogens (GPP) list \citep{organization_prioritization_2017} to guide discovery, research and development of new antibiotics for drug-resistant bacterial infections. Besides tuberculosis, the global priority due to being the most common and lethal airborne AMR disease worldwide today, responsible for 250 000 deaths each year, it includes 12 groups of pathogens in three priority categories:

\begin{itemize}
  \item{\textbf{Critical priority}}
  \begin{itemize}
    \item{carbapenem-resistant \textit{Acinetobacter baumannii}}
    \item{carbapenem-resistant \textit{Pseudomonas aeruginosa}}
    \item{carbapenem-resistant, third generation cephalosporin-resistant \textit{Enterobacteriaceae}} 
  \end{itemize}
  \item{\textbf{High priority}}
  \begin{itemize}
    \item{vancomycin-resistant \textit{Enterococcus faecium}}
    \item{vancomycin-resistant, methicillin-resistant \textit{Staphylococcus aureus}}
    \item{clarithromycin-resistant \textit{Helicobacter pylori}}
    \item{fluoroquinolone-resistant \textit{Campylobacter} species}
    \item{fluoroquinolone-resistant \textit{Salmonella} species}
    \item{fluoroquinolone-resistant, third generation cephalosporin-resistant, \textit{Neisseria gonorrhoeae}}
  \end{itemize}
  \item{\textbf{Medium priority}}
    \begin{itemize}
    \item{penicillin-non-susceptible \textit{Streptococcus pneumoniae}}
    \item{ampicillin-resistant \textit{Haemophilus influenzae}}
    \item{fluoroquinolone-resistant \textit{Shigella} species} 
  \end{itemize}
\end{itemize}

Clinical microbiology is a discipline focused on rapidly characterising pathogen samples in order to direct the management of individual infected patients (diagnostic microbiology) and monitor the epidemiology of infectious disease (public health microbiology), including the detection of outbreaks and infection prevention. Globally, ... FINANTIAL BURDEN!

\subsection{Current standards for diagnostic and public health in clinical microbiology}

The past few decades have seen a major revolution in the operation of microbial laboratories, driven by the development of molecular technologies and ways to make these accessible.

For patients with bacterial infections, the crucial steps are (1) to grow an isolate from a specimen, (2) identify its species, and (3) determine its pathogenic potential and test its susceptibility to antimicrobial drugs. Together this information facilitates the specific and rational treatment of patients. For public health purposes, knowledge also needs to be gained about (4) the relatedness of the pathogen to other strains of the same species to investigate transmission routes and enable the recognition of outbreaks. Traditionally, all of these steps rely on the isolation of the pathogen through culture, followed by a multi-step process that can take days to weeks before obtaining results.  \citep{didelot_transforming_2012}

The traditional approaches to laboratory diagnosis of viral infections have been (1) direct detection in patient material of virions, viral antigens, or viral nucleic acids, (2) isolation of virus in cultured cells, followed by identification of the isolate, and (3) detection and measurement of antibodies in the patient’s serum (serology). \citep{burrell_laboratory_2017}

Currently, the detection and diagnosis of parasite infections rely on several laboratory methods in addition to clinical symptoms, clinical history, travel history, and geographic location of patient. serology-based assays, MOLECULAR-BASED ASSAYS	, PROTEOMICS, microscopy  \citep{ndao_diagnosis_2009}
    %\item The plethora of new antiviral agents and a more sophisticated understanding of how these should be used;
    %\item An increased awareness of the clinical value of, and demand for, prompt information about viral loads, viral sequence data, and antiviral resistance information.

\section{A genomic approach to clinical microbiology}

Since the publication of the first complete microbial genome a quarter of a century ago, that of the bacterium \textit{Haemophilus influenzae} \citep{hood_dna_1996}, genomics has transformed the field of microbiology, and in particular its clinical application.

\subsection{Sequencing}

@@

\section{Metagenomics}

### 2022-01-28 main.tex
update
Progress on introduction

%-----------------------------------------------------------------
% Preamble
%------------------------------------------------------------------
\documentclass[12pt,a4paper,twoside,openright]{book}
\usepackage{mathptmx} % times no roman-like font
\usepackage[left=30mm, right=25mm, top=25mm, bottom=25mm]{geometry} % margins
\pagestyle{fancy}
\renewcommand{\baselinestretch}{1.15}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{figures/}}

%-----------------------------------------------------------------
% Document Metadata
%------------------------------------------------------------------
\title{Thesis Title}

%-----------------------------------------------------------------
% Packages
%------------------------------------------------------------------
\usepackage{setspace}
\usepackage{fancyhdr}
\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\thechapter. #1 }}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\fancyhf{}
\fancyhead[RO]{\bfseries\rightmark}
\fancyhead[LE]{\bfseries\leftmark}
\setlength{\headheight}{15pt}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\addtolength{\headheight}{0.5pt}
\fancypagestyle{plain}{
  \fancyhead{}
  \renewcommand{\headrulewidth}{0pt}
}
\usepackage{hyperref} % To add urls
\usepackage{graphicx} % To add imagess
% Comentar para não incluir a lista de figuras e tabelas no índice
\usepackage[nottoc]{tocbibind}
\usepackage[backend=biber,style=authoryear,maxcitenames=1,autocite=inline,natbib=true]{biblatex} % References
\addbibresource{references.bib}
\usepackage{indentfirst} % Add identation to first paragraph

% Incluir até à subsecção no índice
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

% Permite remover os espaços entre itens
\usepackage{enumitem}

% Caption para imagens
\usepackage[font=footnotesize]{caption}

% Referenciar anexos
\usepackage[toc,page]{appendix}
\usepackage{pdfpages}

%Alterar nome de capitulo para a referencia
\usepackage{fmtcount}

\usepackage{url}

%Para expressões Matemáticas
\usepackage{amsmath}
\usepackage{bm}

%Bibliotecas
%\usepackage{biblatex}
%\usepackage[portuguese]{babel}
\usepackage[utf8]{inputenc}
\usepackage{filecontents}
\usepackage{float}
\raggedbottom
\usepackage{subcaption}
\usepackage{eurosym}
\usepackage{booktabs}
\usepackage[T1]{fontenc}

@@

\begin{document}
\begin{sloppy}

\frontmatter
%-----------------------------------------------------------------
% Cover
%------------------------------------------------------------------

\input{titlepage}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

@@

%-----------------------------------------------------------------
% Dedication
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\vspace*{3cm}
\begin{center}
\emph{``I'm so happy to be done with this shit''} \\
\end{center} 
\begin{flushright}- Inês Mendes\end{flushright}

%-----------------------------------------------------------------
% Acknowledgements
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Acknowledgments}
\chapter*{Acknowledgements}

@@

\keywords{one, two, three, four, five}

@@

\keywords{um, dois, três, quatro, cinco}

%-----------------------------------------------------------------
% Thesis Outline
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Thesis Outline}
\chapter*{Thesis Outline}

%-----------------------------------------------------------------
% Abbreviations
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Abbreviation}
\chapter*{Abbreviation}

%-----------------------------------------------------------------
% Table of Contents
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Table of Contents}
\renewcommand{\contentsname}{Table of Contents}
\tableofcontents

@@

%-----------------------------------------------------------------
% Introduction
%------------------------------------------------------------------
\newpage
\chapter{Introduction\label{ch:introduction}}
\input{chapters/introduction}

%-----------------------------------------------------------------
% Conclusion
%-----------------------------------------------------------------
\newpage
\chapter{Conclusion \label{ch:conclusion}}
\input{chapters/conclusion}

%-----------------------------------------------------------------
% Appendix
%-----------------------------------------------------------------
\appendix
\chapter{Appendix}
\input{chapters/appendix}
\includepdf[pages={1-}]{resources/papers/s41598-018-31873-w.pdf}
\includepdf[pages={1-}]{resources/papers/dkz363.pdf}
\includepdf[pages={1-}]{resources/papers/mgen000328.pdf}

>>>>

%-----------------------------------------------------------------
% Preamble
%------------------------------------------------------------------
\documentclass[12pt,a4paper,twoside,openright]{book}
\usepackage{mathptmx} % times no roman-like font
\usepackage[left=30mm, right=25mm, top=25mm, bottom=25mm]{geometry} % margins
\pagestyle{fancy}
\renewcommand{\baselinestretch}{1.15}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{figures/}}
\def\blankpage{%
      \clearpage%
      \thispagestyle{empty}%
      \null%
      \clearpage}

%-----------------------------------------------------------------
% Packages
%------------------------------------------------------------------
\usepackage{setspace}
\usepackage{fancyhdr}
\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\thechapter. #1 }}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\usepackage{titlesec}
\fancyhf{}
\fancyhead[RO]{\bfseries\rightmark}
\fancyhead[LE]{\bfseries\leftmark}
\setlength{\headheight}{15pt}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\addtolength{\headheight}{0.5pt}
\fancypagestyle{plain}{
  \fancyhead{}
  \renewcommand{\headrulewidth}{0pt}
}
\usepackage{hyperref} % To add urls
\usepackage{graphicx} % To add images
\usepackage[nottoc]{tocbibind}
\usepackage[backend=biber,style=authoryear,maxcitenames=1,autocite=inline,natbib=true,sorting=nty]{biblatex} % References
\addbibresource{references.bib}
\usepackage{indentfirst} % Add indentation to first paragraph
\setlength{\parskip}{1em}
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}
\usepackage{enumitem}
\usepackage[font=footnotesize]{caption}
\usepackage[toc,page]{appendix}
\usepackage{pdfpages}
\usepackage{fmtcount}
\usepackage{url}
\usepackage{amsmath}
\usepackage{bm}
\usepackage[utf8]{inputenc}
\usepackage{filecontents}
\usepackage{float}
\raggedbottom
\usepackage{subcaption}
\usepackage{eurosym}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[acronym]

@@

\begin{document}
\begin{sloppy}
\frontmatter

%-----------------------------------------------------------------
% Cover
%------------------------------------------------------------------
\input{titlepage}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

@@

%-----------------------------------------------------------------
% Disclaimer
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\vspace*{\fill}
As opiniões expressas nesta publicação são da exclusiva responsabilidade do seu autor.
\blankpage

%-----------------------------------------------------------------
% Dedication
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\vspace*{3cm}
\begin{center}
\emph{``I'm so happy to be done with this shit''} \\
\end{center} 
\begin{flushright}- Inês Mendes\end{flushright}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Acknowledgements
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Acknowledgements}
\chapter*{Acknowledgements}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

@@

\keywords{one, two, three, four, five}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

@@

\keywords{um, dois, três, quatro, cinco}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Thesis Outline
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Thesis Outline}
\chapter*{Thesis Outline}
\input{chapters/thesis_outline}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Abbreviations
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Abbreviation}
\chapter*{Abbreviation}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Table of Contents
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Table of Contents}
\renewcommand{\contentsname}{Table of Contents}
\tableofcontents
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\thispagestyle{plain}
\listoftables
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\thispagestyle{plain}
\listoffigures
\clearpage \thispagestyle{empty}\mbox{}\clearpage

@@

%-----------------------------------------------------------------
% Introduction
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{General Introduction\label{ch:introduction}}
\thispagestyle{empty}
\newpage
\begin{refsection}
\include{chapters/introduction}
\newpage
\printbibliography
\end{refsection}

%-----------------------------------------------------------------
% Conclusion
%-----------------------------------------------------------------
\newpage
\chapter{Conclusion \label{ch:conclusion}}
\thispagestyle{empty}
\newpage
\include{chapters/conclusion}

%-----------------------------------------------------------------
% Appendix
%-----------------------------------------------------------------
\appendix
\chapter{Appendix}
\include{chapters/appendix}
\includepdf[pages={1-}]{resources/papers/s41598-018-31873-w.pdf}
\includepdf[pages={1-}]{resources/papers/dkz363.pdf}
\includepdf[pages={1-}]{resources/papers/mgen000328.pdf}
